The Country-level Impact of Educational Patterns on the Rate of Entrepreneurship

### 2014-10-24 EducProposal.Rmd
Update EducProposal.Rmd
Minor formatting and bib edits

# Install necessary packages
# Install rempis from Christpher Gandrud's website
devtools::install_github('christophergandrud/repmis')
library(repmis)
# Load citation package
pkgs <- c('dplyr','ggplot2','repmis')
repmis::LoadandCite(pkgs, file= 'RpackageCitations.bib')

@@

When it comes to public budgeting for education policies, the decision process often does not appear to be rational. That is, it does not seem to take into account the return on investment that should be expected in the long run. Although many claims have been made in the entrepreneurship literature that start-up activities at a country-level are affected by factors other than structural and economic, including cultural and educational. As a matter of fact, only a few empirical studies have actually been carried out and most of them focused on which attitudes were favourable to starting a business, while almost none explored how cultural patterns affected these dispositions.

When cultural factors are examined, they are considered as a given feature that can hardly be traced. By having a closer look at education systems, we intend to explore the processes through which some of these cultural patterns are enhanced and passed from one generation to the next. Aside from parental education, schools are indeed the very place where children are taught how to fit into society. This research is an attempt at looking at one trait in particular, self-confidence.

@@

Drawing a comparison between France and Britain, to what extent do the behavioural characteristics of the education system - individualistic, group-oriented, more or less positive reinforcement in teaching - affect the rate of entrepreneurship in each country?

@@

What determines entrepreneurship at a country-level has been long searched already but although economic, demographic or geographic factors have been looked at extensively, much less has been done regarding cultural or educational factors. However, as Freytag and Thurik stated in 2007, “the relative stability of differences in entrepreneurship across countries suggests that other than economic factors are at play” [@Freytag2007] . Even if “in recent years research has increasingly devoted itself to the subject” [@OECD2009], a lot still needs to be done as the few studies that explored the relationship between attitudes and foundation activities barely searched what determined these attitudes.

Over a hundred years have passed since Max Weber tried to explicit how religious-ethical motivations affect entrepreneurship (Max Weber, 1905), and religion may be less effective in justifying why some are more likely to start businesses than others. Still, that was a founding stone and it is now widely admitted that “cultural features influence attitudes towards start-ups and that these attitudes, in turn, have an impact on start-up activities” [@OECD2009].

@@

This thesis is an attempt at finding out if different patterns in education lead to different outcomes in terms of start-up activities. Indeed, like many in the education literature tend to assert, we believe there is an effect of school personnel on student self-esteem [@Scott1996]. We will explore further the way this mechanism works. As for entrepreneurship literature, it generally acknowledges the influence of self-confidence on start-up activities although “self-confidence in one’s own capabilities to start a business has been found to depend almost exclusively on the individual features of the respective person and his or her integration into social networks” [@OECD2009]. We intend to show that these individual characteristics are partly shaped at an institutional and collective level through the education systems and staff attitudes.

Of course, King and Sobel [@Kerry2008], among others, have already tried to show how school choice can increase the rate of youth entrepreneurship but their perspective was slightly different. They  asserted that more business-like environment in K-12 education was the key, which was favoured by voucher programmes. We believe there is a more widely spread systematic effect of cultural and educational patterns. David Watkins [@Watkins2010] cross-cultural perspective on teaching and learning shall be useful in that respect. Kai-Ming Cheng and Kam-Cheung Wong’s study about school effectiveness in East Asia [@Cheng2006] could also be of help.

Also, as self-confidence is considered to rely partly on one’s integration within social networks [@OECD2009], we will refer to cross-cultural studies such as that of Green, Deschamps and Páez [@green2005variation] about individualism and collectivism, what Freeman [@freeman1997demographic] called idiocentrism and allocentrism. Raeff, Greenfield and Quiroz [@raeff2000conceptualizing] have also conceptualised interpersonal relationships in both contexts. More school-focused, Yamauchi [@yamauchi1998individualism] as well as Davies and Aurini [@davies2003homeschooling] have looked at individualism and collectivism in the classroom and in pedagogy respectively. All these will be precious on the way.

@@

From GEM, we will use the most recent data available regarding the fear of failure rate and perceived capabilities).

@@

5. Return on Investment: Future benefits of a euro spent at present. Here return on investment would only start to be visible after some time, once most of a generation has got into active life which may take up to five to ten years. But even then, not all the expected propensity to start a business will be palpable. When you add ot it the mechanism of the social discount rate, that huge distance in time as of now seem to prevent any calculations of return on investment when it comes to public budgeting for education policies.

>>>>

# Install necessary packages
# Install rempis from Christpher Gandrud's website
devtools::install_github('christophergandrud/repmis')
library(repmis)
# Load citation package
pkgs <- c('repmis')
repmis::LoadandCite(pkgs, file= 'RpackageCitations.bib')

@@

The process of allocating public funds for education is an opportunity to make calculated long-term investments in a nation's populace.  The rationale for decisions made during this process, however, is often difficult to understand.  For example, while many claims have been made in entrepreneurship literature regarding the importance of cultural and educational factors -- not just structural and economic factors -- in determining individual and societal tendencies to start a business, these factors are often not included in the decisionmaking process. As a matter of fact, only a few empirical studies have actually been carried out on this topic and most of them focused on which attitudes were favourable to starting a business, while almost none explored how cultural patterns affected these dispositions.

When cultural factors are examined, they are considered as a given feature that can hardly be traced. By having a closer look at education systems, however, we intend to explore the processes through which some of these cultural patterns are enhanced and passed from one generation to the next. Aside from parental education, schools are indeed the very place where children are taught how to fit into society. This research is an attempt to analyze  one trait in particular, self-confidence.

@@

Drawing a comparison between France and Britain, to what extent do the behavioural characteristics of the education system - individualistic, group-oriented, and more or less positive reinforcement in teaching - affect the rate of entrepreneurship in each country?

@@

What determines entrepreneurship at a country-level has been long searched already but although economic, demographic or geographic factors have been looked at extensively, much less has been done regarding cultural or educational factors. However, as Freytag and Thurik stated in 2007, “the relative stability of differences in entrepreneurship across countries suggests that factors other than economic ones are at play” [@Freytag2007] . Even if “in recent years research has increasingly devoted itself to the subject” [@OECD2009], a lot still needs to be done, as the few studies that explored the relationship between attitudes and foundation activities barely searched what determined these attitudes.

Over a hundred years have passed since Max Weber tried to explain how religious-ethical motivations affect entrepreneurship [@weberprotestant], and religion may be less effective in justifying why some are more likely to start businesses than others. Still, that was a founding stone and it is now widely admitted that “cultural features influence attitudes towards start-ups and that these attitudes, in turn, have an impact on start-up activities” [@OECD2009].

@@

This thesis is an attempt at finding out if different patterns in education lead to different outcomes in terms of start-up activities. Indeed, like many in the education literature tend to assert, we believe there is an effect by school personnel on student self-esteem [@Scott1996]. We will explore further the way this mechanism works. As for entrepreneurship literature, it generally acknowledges the influence of self-confidence on start-up activities although “self-confidence in one’s own capabilities to start a business has been found to depend almost exclusively on the individual features of the respective person and his or her integration into social networks” [@OECD2009]. We intend to show that these individual characteristics are partly shaped at an institutional and collective level through the education systems and staff attitudes.

Of course, King and Sobel [@Kerry2008], among others, have already tried to show how school choice can increase the rate of youth entrepreneurship, but their perspective was slightly different. They  asserted that a more business-like environment in K-12 education was the key, which was favoured by voucher programmes. We believe there is a more widely spread systematic effect of cultural and educational patterns. David Watkins [@Watkins2010] cross-cultural perspective on teaching and learning shall be useful in that respect. Kai-Ming Cheng and Kam-Cheung Wong’s study about school effectiveness in East Asia [@Cheng2006] could also be of help.

Also, as self-confidence is considered to rely partly on one’s integration within social networks [@OECD2009], we will refer to cross-cultural studies such as that of Green, Deschamps and Páez [@green2005variation] about individualism and collectivism, what Freeman [@freeman1997demographic] called idiocentrism and allocentrism. Raeff, Greenfield and Quiroz [@raeff2000conceptualizing] have also conceptualised interpersonal relationships in both contexts. More school-focused, Yamauchi [@yamauchi1998individualism] as well as Davies and Aurini [@davies2003homeschooling] have looked at individualism and collectivism in the classroom and in pedagogy respectively. All these will be important on the way.

@@

From GEM, we will use the most recent data available regarding the fear of failure rate and perceived capabilities.

@@

5. Return on Investment: Future benefits of a euro spent at present. Here return on investment would only start to be visible after some time, once most of a generation has got into active life which may take up to five to ten years. But even then, not all the expected propensity to start a business will be palpable. When you add it to the mechanism of the social discount rate, that huge distance in time as of now seem to prevent any calculations of return on investment when it comes to public budgeting for education policies.
